\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-10-23 doc/flexiblesusy-paper.tex
Fix a typo in the paper

\newpage
\section{Introduction}
Supersymmetry (SUSY) provides the only non-trivial way to extend the
space-time symmetries of the Poincar\'e
group, which still has scattering in the resulting quantum theory \cite{Coleman:1967ad,Haag:1974qh}. This leads many to suspect that
SUSY may be realized in nature in some form. In particular
supersymmetric extensions of the standard model (SM) where SUSY is broken
at the TeV scale have been proposed to solve the hierarchy
problem \cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
  Susskind:1978ms, 'tHooft:1980xb}, allow gauge coupling
unification \cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
  Langacker:1991an, Giunti:1991ta} and predict a dark matter candidate
which can fit the observed relic
density \cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
been used for baryogenesis or leptogensis to solve the
matter-anti-matter asymmetry of the universe and have been considered
as the low energy effective models originating from string
theory.

Detailed phenomenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM).  Such work
has been greatly aided by public spectrum generators for the MSSM
\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
allowing fast and reliable exploration of the sparticle spectrum,
mixings and couplings, which can be obtained from particular choices
of breaking mechanism inspired boundary conditions and specified
parameters. Beyond the MSSM there are also two public spectrum
generators \cite{Ellwanger:2006rn,Allanach:2013kza} for the next to
minimal supersymmetric standard model (NMSSM) \cite{NMSSM} (or for recent
reviews see \cite{Ellwanger:2009dp,Maniatis:2009re}).

@@

These developments motivate the exploration of non-minimal SUSY models
which ameliorate the naturalness problems, as can happen in the \ESSM
\cite{King:2005jy,King:2005my,King:2007uj,Athron:2010zz,Nevzorov:2012hs,Nevzorov:2013tta,Nevzorov:2013ixa}, USSM
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq} or from other gauge extensions \cite{Batra:2003nj, Medina:2009ey, Bharucha:2013ela}, 
by raising the tree level Higgs mass. At the same time they can also
motivate models that are developed with a fresh perspective, based on
other considerations.  In both cases exploration of such models can be
aided if it is possible to quickly create a fast spectrum generator.
Currently there is only one option for this, a SPheno-like FORTRAN
code which can be generated from \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.

@@

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the NMSSM \cite{NMSSM} in it's $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM})
\cite{Fayet:1977yc,Suematsu:1994qm,Cvetic:1995rj,deCarlos:1997yv,Cvetic:1997ky,Demir:1998dk,Langacker:1998tc,Erler:2002pr,Choi:2006fz,Ham:2007wc,Langacker:2008yv,Ham:2008xf,Kalinowski:2008iq},
the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
predefined model files.  For all these models spectrum generators can
be generated easily like for the CMSSM in
\secref{sec:quick-start-cmssm}.  The spectrum generator for the
$Z_3$-symmetric NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
One of the design goals is modularity and the possibility to easily
construct custom spectrum generators.  The details of the
customization can be found in Sections
\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
the NMSSM.  The NMSSM variant above unifies all soft-breaking
trilinear scalar couplings at the GUT scale.  In order to relax this
constraint and use a separate value for $A_\lambda$ at the GUT scale
one can edit the model file \code{model_files/NMSSM/FlexibleSUSY.m.in} and
change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

SUSYScale = Sqrt[M[Su[1]]*M[Su[6]]];

@@

  SUSY-scale, $M_S$, is usually defined as $M_S =
  \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$.  However, in the example
  above, where sfermion flavour violation is enabled, it has the value
  $M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}}$, where
  $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the \DRbar\ masses of
  the lightest and heaviest up-type squark, respectively.

@@

\fs allows the user to add leading two-loop contributions to the
CP-even Higgs tadpoles and self-energies.  For MSSM-like models (with
two CP-even Higgs bosons, one CP-odd Higgs boson, one neutral
Goldstone boson) routines dor caluclating these corrections will be
generated by setting \code{UseHiggs2LoopMSSM = True} in the model file
and by defining the effective $\mu$-term \code{EffectiveMu = \\[Mu]}.
This will add the zero-momentum corrections of the order $O(y_t^4 +
y_b^2 y_t^2 + y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$,
$O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are generated by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq} and also the MSSM contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
\cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
correction for that order in the NMSSM, but can be a good approximation
when singlet mixing is very small\footnote{These corrections may be disabled in the SLHA file, as described in section \secref{sec:PoleMasses}.}.

@@

\item The \emph{SUSY-scale constraint} is intended to set boundary
  conditions at the mass scale $M_S$ of the SUSY particles.  The value
  of $M_S$ is defined by the model file variable \code{SUSYScale}.  In
  the example model file in \secref{sec:modfile} it is defined to be
  $\sqrt{m_{\tilde{u}_1} m_{\tilde{u}_6}}$, where $m_{\tilde{u}_1}$ is
  the mass of the lightest, and $m_{\tilde{u}_6}$ is the mass of the
  heaviest up-type squark.  The \code{apply()} function for this
  constraint sets the model parameters to the values defined in
  \code{SUSYScaleInput}.  Afterwards, \code{apply()} solves the EWSB
  equations at the loop level by adjusting the parameters given in
  \code{EWSBOutputParameters} such that the effective Higgs potential
  is minimized.  See \secref{sec:ewsb} for a more detailed description
  of the EWSB in \fs.

@@

      M_S = \sqrt{m_{\tilde{u}_1}m_{\tilde{u}_6}},

@@

    where $m_{\tilde{u}_1}$ and $m_{\tilde{u}_6}$ are the lightest and
    heaviest up-type squarks, respectively.

@@

\fs generated NUHM \ESSM has also been compared against a handwritten
spectrum generator for a constrained version of the \ESSM
\cite{Athron:2009ue, Athron:2009bs, Athron:2011wu, Athron:2012sq,
  Athron:2012pw, Athron:2013ipa}.  The RGEs could be compared directly
in unit tests and were found to match.  The handwritten code doesn't
include full one-loop self-energies or tadpoles so tests on these were
not carried out.  Although the generators assume different constraints
they could be compared by using the output of the CE$_6$SSM generator
as an input to \fs and the spectra were found to be in reasonable
agreement, given the different levels of precision.

@@

%\cite{Girardello:1981wz}
\bibitem{Girardello:1981wz}
  L.~Girardello and M.~T.~Grisaru,
  %``Soft Breaking of Supersymmetry,''
  Nucl.\ Phys.\ B {\bf 194} (1982) 65.
  %%CITATION = NUPHA,B194,65;%%
  %717 citations counted in INSPIRE as of 21 Feb 2014

@@

%\cite{Ellwanger:2006rn}
\bibitem{Ellwanger:2006rn} 
  U.~Ellwanger and C.~Hugonie,
  %``NMSPEC: A Fortran code for the sparticle and Higgs masses in the NMSSM with GUT scale boundary conditions,''
  Comput.\ Phys.\ Commun.\  {\bf 177}, 399 (2007)
  [hep-ph/0612134].
  %%CITATION = HEP-PH/0612134;%%
  %81 citations counted in INSPIRE as of 03 Nov 2013
%\cite{Allanach:2013kza}
\bibitem{Allanach:2013kza} 
  B.~C.~Allanach, P.~Athron, L.~C.~Tunstall, A.~Voigt and A.~G.~Williams,
  %``Next-to-Minimal SOFTSUSY,''
  Comput.\ Phys.\ Commun.\  {\bf 185}, 2322 (2014)
  [arXiv:1311.7659 [hep-ph]].
  %%CITATION = ARXIV:1311.7659;%%
  %2 citations counted in INSPIRE as of 15 Jun 2014
\bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
B \textbf{64} (1976) 159; Phys. Lett. B \textbf{69} (1977) 489 and Phys. Lett. B
\textbf{84} (1979) 416; H.P. Nilles, M. Srednicki and D. Wyler, Phys. Lett. B
\textbf{120} (1983) 346; J.M. Frere, D.R. Jones and S. Raby, Nucl. Phys. B
\textbf{222} (1983) 11; J.P. Derendinger and C.A. Savoy, Nucl. Phys. B
\textbf{237} (1984) 307;  A.I. Veselov, M.I. Vysotsky and K.A. Ter-Martirosian,
Sov. Phys. JETP \textbf{63} (1986) 489; J.R. Ellis, J.F. Gunion, H.E. Haber, L.
Roszkowski and F. Zwirner, Phys. Rev. D \textbf{39}  (1989) 844; M. Drees, Int.
J. Mod. Phys. A \textbf{4}  (1989) 3635; U. Ellwanger, M. Rausch de
Traubenberg and C.A. Savoy, Phys. 
Lett. B \textbf{315} (1993) 331, Z. Phys. C {\bf 67} (1995) 665 and Nucl. Phys.
B \textbf{492} (1997) 307; U.~Ellwanger, Phys.\ Lett.\  B {\bf 303} (1993) 271; P.
Pandita, Z. Phys. C \textbf{59} (1993) 575; T. Elliott, S.F. King and P.L.
White, Phys. Rev. D {\bf 49} (1994) 2435; S.F. King and P.L. White, Phys. Rev. D
\textbf{52} (1995) 4183;  F.~Franke and H.~Fraas, Int.\ J.\ Mod.\ Phys.\  A {\bf
12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
%\cite{Ellwanger:2009dp}
\bibitem{Ellwanger:2009dp} 
  U.~Ellwanger, C.~Hugonie and A.~M.~Teixeira,
  %``The Next-to-Minimal Supersymmetric Standard Model,''
  Phys.\ Rept.\  {\bf 496}, 1 (2010)
  [arXiv:0910.1785 [hep-ph]].
  %%CITATION = ARXIV:0910.1785;%%
  %327 citations counted in INSPIRE as of 21 Sep 2013
%\cite{Maniatis:2009re}
\bibitem{Maniatis:2009re} 
  M.~Maniatis,
  %``The Next-to-Minimal Supersymmetric extension of the Standard Model reviewed,''
  Int.\ J.\ Mod.\ Phys.\ A {\bf 25}, 3505 (2010)
  [arXiv:0906.0777 [hep-ph]].
  %%CITATION = ARXIV:0906.0777;%%
  %160 citations counted in INSPIRE as of 09 Jun 2014

@@

@@

%\cite{King:2005my}
\bibitem{King:2005my} 
  S.~F.~King, S.~Moretti and R.~Nevzorov,
  %``Exceptional supersymmetric standard model,''
  Phys.\ Lett.\ B {\bf 634}, 278 (2006)
  [hep-ph/0511256].
  %%CITATION = HEP-PH/0511256;%%
  %90 citations counted in INSPIRE as of 09 Jun 2014

%\cite{King:2007uj}
\bibitem{King:2007uj} 
  S.~F.~King, S.~Moretti and R.~Nevzorov,
  %``Gauge coupling unification in the exceptional supersymmetric standard model,''
  Phys.\ Lett.\ B {\bf 650}, 57 (2007)
  [hep-ph/0701064].
  %%CITATION = HEP-PH/0701064;%%
  %46 citations counted in INSPIRE as of 09 Jun 2014

@@

%\cite{Nevzorov:2013tta}
\bibitem{Nevzorov:2013tta} 
  R.~Nevzorov and S.~Pakvasa,
  %``Exotic Higgs decays in the $E_6$ inspired SUSY models,''
  Phys.\ Lett.\ B {\bf 728}, 210 (2014)
  [arXiv:1308.1021 [hep-ph]].
  %%CITATION = ARXIV:1308.1021;%%
  %2 citations counted in INSPIRE as of 13 Jun 2014

%\cite{Nevzorov:2013ixa}
\bibitem{Nevzorov:2013ixa} 
  R.~Nevzorov,
  %``Quasi-fixed point scenarios and the Higgs mass in the E6 inspired SUSY models,''
  Phys.\ Rev.\ D {\bf 89}, 055010 (2014)
  [arXiv:1309.4738 [hep-ph]].
  %%CITATION = ARXIV:1309.4738;%%
  %1 citations counted in INSPIRE as of 13 Jun 2014

%\cite{Fayet:1977yc}
\bibitem{Fayet:1977yc}
  P.~Fayet,
  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
  Phys.\ Lett.\ B {\bf 69} (1977) 489.
  %%CITATION = PHLTA,B69,489;%%
  %817 citations counted in INSPIRE as of 04 Jun 2014
%\cite{deCarlos:1997yv}
%\cite{Suematsu:1994qm}
\bibitem{Suematsu:1994qm} 
  D.~Suematsu and Y.~Yamagishi,
  %``Radiative symmetry breaking in a supersymmetric model with an extra U(1),''
  Int.\ J.\ Mod.\ Phys.\ A {\bf 10}, 4521 (1995)
  [hep-ph/9411239].
  %%CITATION = HEP-PH/9411239;%%
  %116 citations counted in INSPIRE as of 09 Jun 2014

  %\cite{Cvetic:1995rj}
\bibitem{Cvetic:1995rj} 
  M.~Cvetic and P.~Langacker,
  %``Implications of Abelian extended gauge structures from string models,''
  Phys.\ Rev.\ D {\bf 54}, 3570 (1996)
  [hep-ph/9511378].
  %%CITATION = HEP-PH/9511378;%%
  %207 citations counted in INSPIRE as of 09 Jun 2014

\bibitem{deCarlos:1997yv} 
  B.~de Carlos and J.~R.~Espinosa,
  %``Cold dark matter candidate in a class of supersymmetric models with an extra U(1),''
  Phys.\ Lett.\ B {\bf 407}, 12 (1997)
  [hep-ph/9705315].
  %%CITATION = HEP-PH/9705315;%%
  %51 citations counted in INSPIRE as of 09 Jun 2014
%\cite{Cvetic:1997ky}
\bibitem{Cvetic:1997ky}
  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
  Phys.\ Rev.\ D {\bf 56} (1997) 2861
   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
  [hep-ph/9703317].
  %%CITATION = HEP-PH/9703317;%%
  %237 citations counted in INSPIRE as of 04 Jun 2014

%\cite{Demir:1998dk}
\bibitem{Demir:1998dk} 
  D.~A.~Demir and N.~K.~Pak,
  %``One loop effects in supergravity models with an additional U(1),''
  Phys.\ Rev.\ D {\bf 57}, 6609 (1998)
  [hep-ph/9809357].
  %%CITATION = HEP-PH/9809357;%%
  %19 citations counted in INSPIRE as of 09 Jun 2014

%\cite{Langacker:1998tc}
\bibitem{Langacker:1998tc} 
  P.~Langacker and J.~Wang,
  %``U(1)-prime symmetry breaking in supersymmetric E(6) models,''
  Phys.\ Rev.\ D {\bf 58}, 115010 (1998)
  [hep-ph/9804428].
  %%CITATION = HEP-PH/9804428;%%
  %108 citations counted in INSPIRE as of 09 Jun 2014

%\cite{Erler:2002pr}
\bibitem{Erler:2002pr}
  J.~Erler, P.~Langacker and T.~-j.~Li,
  %``The $Z$ - $Z^\prime$ mass hierarchy in a supersymmetric model with a secluded U(1) -prime breaking sector,''
  Phys.\ Rev.\ D {\bf 66} (2002) 015002
  [hep-ph/0205001].
  %%CITATION = HEP-PH/0205001;%%
  %84 citations counted in INSPIRE as of 04 Jun 2014

%\cite{Choi:2006fz}
\bibitem{Choi:2006fz} 
  S.~Y.~Choi, H.~E.~Haber, J.~Kalinowski and P.~M.~Zerwas,
  %``The Neutralino sector in the U(1)-extended supersymmetric standard model,''
  Nucl.\ Phys.\ B {\bf 778}, 85 (2007)
  [hep-ph/0612218].
  %%CITATION = HEP-PH/0612218;%%
  %43 citations counted in INSPIRE as of 09 Jun 2014

%\cite{Ham:2007wc}
\bibitem{Ham:2007wc} 
  S.~W.~Ham and S.~K.~OH,
  %``Electroweak phase transition in MSSM with U(1)-prime in explicit CP violation scenario,''
  Phys.\ Rev.\ D {\bf 76}, 095018 (2007)
  [arXiv:0708.1785 [hep-ph]].
  %%CITATION = ARXIV:0708.1785;%%
  %20 citations counted in INSPIRE as of 09 Jun 2014

%\cite{Ham:2008xf}
\bibitem{Ham:2008xf} 
  S.~W.~Ham, T.~Hur, P.~Ko and S.~K.~Oh,
  %``Neutral scalar Higgs bosons in the USSM at the LHC,''
  J.\ Phys.\ G {\bf 35}, 095007 (2008)
  [arXiv:0801.2361 [hep-ph]].
  %%CITATION = ARXIV:0801.2361;%%
  %10 citations counted in INSPIRE as of 09 Jun 2014

%\cite{Kalinowski:2008iq}
\bibitem{Kalinowski:2008iq} 
  J.~Kalinowski, S.~F.~King and J.~P.~Roberts,
  %``Neutralino Dark Matter in the USSM,''
  JHEP {\bf 0901}, 066 (2009)
  [arXiv:0811.2204 [hep-ph]].
  %%CITATION = ARXIV:0811.2204;%%
  %30 citations counted in INSPIRE as of 09 Jun 2014
%\cite{Bharucha:2013ela}

%\cite{Batra:2003nj}
\bibitem{Batra:2003nj} 
  P.~Batra, A.~Delgado, D.~E.~Kaplan and T.~M.~P.~Tait,
  %``The Higgs mass bound in gauge extensions of the minimal supersymmetric standard model,''
  JHEP {\bf 0402}, 043 (2004)
  [hep-ph/0309149].
  %%CITATION = HEP-PH/0309149;%%
  %174 citations counted in INSPIRE as of 12 Jun 2014
%\cite{Medina:2009ey}
\bibitem{Medina:2009ey} 
  A.~D.~Medina, N.~R.~Shah and C.~E.~M.~Wagner,
  %``A Heavy Higgs and a Light Sneutrino NLSP in the MSSM with Enhanced SU(2) D-terms,''
  Phys.\ Rev.\ D {\bf 80}, 015001 (2009)
  [arXiv:0904.1625 [hep-ph]].
  %%CITATION = ARXIV:0904.1625;%%
  %9 citations counted in INSPIRE as of 12 Jun 2014

@@

%\cite{Athron:2011wu}
\bibitem{Athron:2011wu} 
  P.~Athron, S.~F.~King, D.~J.~Miller, S.~Moretti and R.~Nevzorov,
  %``LHC Signatures of the Constrained Exceptional Supersymmetric Standard Model,''
  Phys.\ Rev.\ D {\bf 84}, 055006 (2011)
  [arXiv:1102.4363 [hep-ph]].
  %%CITATION = ARXIV:1102.4363;%%
  %25 citations counted in INSPIRE as of 09 Jun 2014

>>>>

\newpage
\section{Introduction}

Supersymmetry (SUSY) provides the only non-trivial way to extend the
space-time symmetries of the Poincar\'e
group, which still has scattering in the resulting quantum theory \cite{Coleman:1967ad,Haag:1974qh}. This leads many to suspect that
SUSY may be realized in nature in some form. In particular
supersymmetric extensions of the standard model (SM) where SUSY is broken
at the TeV scale have been proposed to solve the hierarchy
problem \cite{Weinberg:1975gm, Weinberg:1979bn, Gildener:1976ai,
  Susskind:1978ms, 'tHooft:1980xb}, allow gauge coupling
unification \cite{Langacker:1990jh, Ellis:1990wk, Amaldi:1991cn,
  Langacker:1991an, Giunti:1991ta} and predict a dark matter candidate
which can fit the observed relic
density \cite{Goldberg:1983nd,Ellis:1983ew}.  Such models have also
been used for baryogenesis or leptogensis to solve the
matter-anti-matter asymmetry of the universe and have been considered
as the low energy effective models originating from string
theory.

Detailed phenomenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM), for a review
see \cite{Chung:2003fi}.  Such work has been greatly aided by public
spectrum generators for the MSSM
\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae,Chowdhury:2011zr},
allowing fast and reliable exploration of the sparticle spectrum,
mixings and couplings, which can be obtained from particular choices
of breaking mechanism inspired boundary conditions and specified
parameters. Beyond the MSSM there are also two public spectrum
generators \cite{Ellwanger:2004xm, Ellwanger:2005dv,
  Ellwanger:2006rn,Ellwanger:2008py, Allanach:2013kza} for the next to
minimal supersymmetric standard model (NMSSM)  (for
recent reviews see \cite{Ellwanger:2009dp, Maniatis:2009re}).

@@

These developments motivate the exploration of non-minimal SUSY models
which ameliorate the naturalness problems by raising the tree level
Higgs mass, as can happen in the USSM
\cite{Fayet:1977yc,Cvetic:1997ky,Langacker:2008yv}, \ESSM
\cite{King:2005jy,Athron:2010zz} and similar models
\cite{Nevzorov:2012hs}, or from other gauge extensions
\cite{Batra:2003nj, Bharucha:2013ela}. At the same time they can also
motivate models that are developed with a fresh perspective, based on
other considerations.  In both cases exploration of such models can be
aided if it is possible to quickly create a fast spectrum generator.
Currently there is only one option for this, a SPheno-like FORTRAN
code which can be generated from \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}.

@@

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the NMSSM in its $Z_3$-symmetric form (called \code{NMSSM}), $Z_3$-violating NMSSM (\code{SMSSM}), the USSM (\code{UMSSM}),
the NUHM \ESSM (\code{E6SSM}) \cite{Athron:2007en}, the
right-handed neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
predefined model files.  For all these models spectrum generators can
be generated easily like for the CMSSM in
\secref{sec:quick-start-cmssm}.  The spectrum generator for the
$Z_3$-symmetric NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
One of the design goals is modularity and the possibility to easily
construct custom spectrum generators.  The details of the
customization can be found in Sections
\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
the NMSSM.  The NMSSM variant above unifies all soft-breaking
trilinear scalar couplings at the GUT scale.  In order to relax this
constraint and use a separate value for $A_\lambda$ at the GUT scale
one can edit the model file \code{model_files/NMSSM/FlexibleSUSY.m.in} and
change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

SUSYScale = Sqrt[Product[M[Su[i]]^(Abs[ZU[i,3]]^2 + Abs[ZU[i,6]]^2), {i,6}]];

@@

  SUSY-scale, $M_S$, is defined as
    M_S = \sqrt{\prod_{i=1}^6 m_{\tilde{u}_i}^{|(Z_u)_{i3}|^2 + |(Z_u)_{i6}|^2}} ,
    \label{eq:definition_of_MS}
  where $m_{\tilde{u}_i}$ is the \DRbar\ mass of the $i$th up-type
  squark and $Z_u$ is the up-type squark mixing matrix.  The
  definition \eqref{eq:definition_of_MS} is equivalent to the usual
  choice $M_S = \sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$ without squark
  flavour mixing.

@@

\fs allows the user to add leading two-loop contributions to the
CP-even and CP-odd Higgs self-energies as well as to the CP-even Higgs
tadpoles.  For MSSM-like models (with two CP-even Higgs bosons, one
CP-odd Higgs boson, one neutral Goldstone boson) routines dor
caluclating these corrections will be generated by setting
\code{UseHiggs2LoopMSSM = True} in the model file and by defining the
effective $\mu$-term \code{EffectiveMu = \\[Mu]}.  This will add the
zero-momentum corrections of the order $O(y_t^4 + y_b^2 y_t^2 +
y_b^4)$, $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$, $O(y_\tau^4)$ from
\cite{Degrassi:2001yf,Brignole:2001jy,Dedes:2002dy,Brignole:2002bz,Dedes:2003km}.
For NMSSM-like models (with three CP-even Higgs bosons, two CP-odd
Higgs bosons, one neutral Goldstone boson) the two-loop contributions
are generated by setting \code{UseHiggs2LoopNMSSM = True} and by
defining the effective $\mu$-term like \code{EffectiveMu = \\[Lambda]
  vS / Sqrt[2]}, for example.  This will add the zero-momentum
corrections of the order $O(y_t^2 g_3^2)$, $O(y_b^2 g_3^2)$ from
\cite{Degrassi:2009yq} and also the MSSM contributions of the order
$O(y_\tau^4)$, $O(y_t^4 + y_t^2 y_b^2 + y_b^4)$ as well
\cite{Brignole:2001jy,Dedes:2003km} which only represent a partial
correction for that order in the NMSSM, but can be a good
approximation when singlet mixing is very small\footnote{These
  corrections may be disabled in the SLHA file, as described in
  section \secref{sec:PoleMasses}.}.

@@

\item The \emph{SUSY-scale constraint} is intended to set boundary
  conditions at the mass scale $M_S$ of the SUSY particles.  The value
  of $M_S$ is defined in the model file variable \code{SUSYScale}.  In
  the example model file in \secref{sec:modfile} it is given by the
  expression written in Eq.~\eqref{eq:definition_of_MS}.  The
  \code{apply()} function for this constraint sets the model
  parameters to the values defined in \code{SUSYScaleInput}.
  Afterwards, \code{apply()} solves the EWSB equations at the loop
  level by adjusting the parameters given in
  \code{EWSBOutputParameters} such that the effective Higgs potential
  is minimized.  See \secref{sec:ewsb} for a more detailed description
  of the EWSB in \fs.

@@

      M_S = \sqrt{\prod_{i=1}^6 m_{\tilde{u}_i}^{|(Z_u)_{i3}|^2 + |(Z_u)_{i6}|^2}} ,

@@

    where $m_{\tilde{u}_i}$ is the \DRbar\ mass of the $i$th up-type
    squark.

@@

\fs generated NUHM \ESSM has also been compared against a handwritten
spectrum generator for a constrained version of the \ESSM
\cite{Athron:2009ue, Athron:2009bs, Athron:2012pw}.  The RGEs could be
compared directly in unit tests and were found to match.  The
handwritten code doesn't include full one-loop self-energies or
tadpoles so tests on these were not carried out.  Although the
generators assume different constraints they could be compared by
using the output of the CE$_6$SSM generator as an input to \fs and the
spectra were found to be in reasonable agreement, given the different
levels of precision.

@@

%\cite{Girardello:1981wz}
\bibitem{Girardello:1981wz}
  L.~Girardello and M.~T.~Grisaru,
  %``Soft Breaking of Supersymmetry,''
  Nucl.\ Phys.\ B {\bf 194} (1982) 65.
  %%CITATION = NUPHA,B194,65;%%
  %717 citations counted in INSPIRE as of 21 Feb 2014
\bibitem{Chung:2003fi} D.~J.~H.~Chung, L.~L.~Everett, G.~L.~Kane,
  S.~F.~King, J.~D.~Lykken and L.~T.~Wang,
  %``The Soft supersymmetry breaking Lagrangian: Theory and applications,''
  Phys.\ Rept.\  {\bf 407}, 1 (2005)
  [hep-ph/0312378].
  %%CITATION = HEP-PH/0312378;%%
  %389 citations counted in INSPIRE as of 23 Oct 2014

@@

%\cite{Ellwanger:2004xm}
\bibitem{Ellwanger:2004xm} 
  U.~Ellwanger, J.~F.~Gunion and C.~Hugonie,
  %``NMHDECAY: A Fortran code for the Higgs masses, couplings and decay widths in the NMSSM,''
  JHEP {\bf 0502}, 066 (2005)
  [hep-ph/0406215].
  %%CITATION = HEP-PH/0406215;%%
  %256 citations counted in INSPIRE as of 23 Oct 2014

%\cite{Ellwanger:2005dv}
\bibitem{Ellwanger:2005dv} 
  U.~Ellwanger and C.~Hugonie,
  %``NMHDECAY 2.0: An Updated program for sparticle masses, Higgs masses, couplings and decay widths in the NMSSM,''
  Comput.\ Phys.\ Commun.\  {\bf 175}, 290 (2006)
  [hep-ph/0508022].
  %%CITATION = HEP-PH/0508022;%%
  %227 citations counted in INSPIRE as of 23 Oct 2014

%\cite{Ellwanger:2006rn}
\bibitem{Ellwanger:2006rn} 
  U.~Ellwanger and C.~Hugonie,
  %``NMSPEC: A Fortran code for the sparticle and Higgs masses in the NMSSM with GUT scale boundary conditions,''
  Comput.\ Phys.\ Commun.\  {\bf 177}, 399 (2007)
  [hep-ph/0612134].
  %%CITATION = HEP-PH/0612134;%%
  %81 citations counted in INSPIRE as of 03 Nov 2013

  %\cite{Ellwanger:2008py}
\bibitem{Ellwanger:2008py} 
  U.~Ellwanger, C.-C.~Jean-Louis and A.~M.~Teixeira,
  %``Phenomenology of the General NMSSM with Gauge Mediated Supersymmetry Breaking,''
  JHEP {\bf 0805}, 044 (2008)
  [arXiv:0803.2962 [hep-ph]].
  %%CITATION = ARXIV:0803.2962;%%
  %34 citations counted in INSPIRE as of 23 Oct 2014

%\cite{Allanach:2013kza}
\bibitem{Allanach:2013kza} 
  B.~C.~Allanach, P.~Athron, L.~C.~Tunstall, A.~Voigt and A.~G.~Williams,
  %``Next-to-Minimal SOFTSUSY,''
  Comput.\ Phys.\ Commun.\  {\bf 185}, 2322 (2014)
  [arXiv:1311.7659 [hep-ph]].
  %%CITATION = ARXIV:1311.7659;%%
  %2 citations counted in INSPIRE as of 15 Jun 2014
%\cite{Ellwanger:2009dp}
\bibitem{Ellwanger:2009dp} 
  U.~Ellwanger, C.~Hugonie and A.~M.~Teixeira,
  %``The Next-to-Minimal Supersymmetric Standard Model,''
  Phys.\ Rept.\  {\bf 496}, 1 (2010)
  [arXiv:0910.1785 [hep-ph]].
  %%CITATION = ARXIV:0910.1785;%%
  %327 citations counted in INSPIRE as of 21 Sep 2013
%\cite{Maniatis:2009re}
\bibitem{Maniatis:2009re} 
  M.~Maniatis,
  %``The Next-to-Minimal Supersymmetric extension of the Standard Model reviewed,''
  Int.\ J.\ Mod.\ Phys.\ A {\bf 25}, 3505 (2010)
  [arXiv:0906.0777 [hep-ph]].
  %%CITATION = ARXIV:0906.0777;%%
  %160 citations counted in INSPIRE as of 09 Jun 2014

@@

%\cite{Fayet:1977yc}
\bibitem{Fayet:1977yc}
  P.~Fayet,
  %``Spontaneously Broken Supersymmetric Theories of Weak, Electromagnetic and Strong Interactions,''
  Phys.\ Lett.\ B {\bf 69} (1977) 489.
  %%CITATION = PHLTA,B69,489;%%
  %817 citations counted in INSPIRE as of 04 Jun 2014
%\cite{deCarlos:1997yv}
%\cite{Suematsu:1994qm}

\bibitem{Cvetic:1997ky}
  M.~Cvetic, D.~A.~Demir, J.~R.~Espinosa, L.~L.~Everett and P.~Langacker,
  %``Electroweak breaking and the mu problem in supergravity models with an additional U(1),''
  Phys.\ Rev.\ D {\bf 56} (1997) 2861
   [Erratum-ibid.\ D {\bf 58} (1998) 119905]
  [hep-ph/9703317].
  %%CITATION = HEP-PH/9703317;%%
  %237 citations counted in INSPIRE as of 04 Jun 2014

@@

@@

%\cite{Batra:2003nj}
\bibitem{Batra:2003nj} 
  P.~Batra, A.~Delgado, D.~E.~Kaplan and T.~M.~P.~Tait,
  %``The Higgs mass bound in gauge extensions of the minimal supersymmetric standard model,''
  JHEP {\bf 0402}, 043 (2004)
  [hep-ph/0309149].
  %%CITATION = HEP-PH/0309149;%%
  %174 citations counted in INSPIRE as of 12 Jun 2014

@@
